---
title: "Medicaid Drug Rebate Program (MDRP) Database"
output: 
  html_notebook:
    code_folding: hide
---

```{r setup, warning=FALSE, message=FALSE}
options(future.globals.maxSize = 2 * 1024 * 10E6);

.cran_libs <- c("purrr", "jsonlite", "httr", "summarytools", "munsell"
                , "SmartEDA", "htmltools", "slider", "stringi", "magrittr"
                , "DT", "data.table", "pdftools", "lubridate", "future", "furrr", "future.callr")
.git_libs <- paste0("book.of.", c("utilities", "features", "workflow")) |> c("architect")
              
if (!"book.of.workflow" %in% dir(sprintf("%s/library", R.home()))){
  if (!"remotes" %in% dir(sprintf("%s/library", R.home()))){ 
    htmltools::tags$span(style = "text-deoration:italic; color:#333333; ", "... installing missing package 'remotes'")
    install.packages("remotes", repos = "https://cloud.r-project.org") 
  }
  
  htmltools::tags$span(
    style = "text-deoration:italic; color:#333333; "
    , "... installing missing packages 'book.of.utilities', 'book.of.features', 'book.of.workflow', 'architect'"
    );
  
  remotes::install_github("delriaan/book.of.utilities", subdir = "pkg")
  remotes::install_github("delriaan/book.of.features", subdir = "pkg")
  remotes::install_github("delriaan/book.of.workflow", subdir = "pkg")
  remotes::install_github("delriaan/architect", subdir = "pkg")
}

library(book.of.workflow)
load_unloaded(!!!.cran_libs, autoinstall = TRUE)
load_unloaded(!!!.git_libs)

urls <- list(
  data = list(
    `Data Dictionary` = "https://www.medicaid.gov/medicaid-chip-program-information/by-topics/prescription-drugs/downloads/recordspecficationanddefinitions.pdf"
    , Data = "https://download.medicaid.gov/data/drugproducts1q_2023.csv"
    ) |> imap(\(x, y) tags$li(tags$a(href = x, y))) |> tags$ol()
  , git_libs = map(.git_libs, \(x) tags$li(tags$a(title = paste0("delriaan/", x), x))) |> list() |> tags$ul()
  , openFDA = "https://download.open.fda.gov/drug/ndc/drug-ndc-0001-of-0001.json.zip"
  )
```

#  {.tabset .tabset-fade .tabset-pills}

## Purpose &<br> Metadata 

This project explores the Medicaid Drug Rebate Program (MDRP) database via [API](https://data.medicaid.gov/dataset/0ad65fe5-3ad3-5d79-a3f9-7893ded7963a) calls (dataset description [here](https://www.medicaid.gov/medicaid/prescription-drugs/medicaid-drug-rebate-program/medicaid-drug-rebate-program-data/index.html)):

`r urls$data`

### Required LIbraries

```{r, echo=FALSE, cache=TRUE, cache.lazy=TRUE}
tags$table(
  style = "width:475px; "
  , tags$tr(
    tags$th(style = "text-align:middle; border: solid 2px #AAAAFF; background-color:#EEEEEE; ", "CRAN", colspan = 2, width="65%")
    , tags$th(style = "text-align:middle; border: solid 2px #AAAAFF; background-color:#AAAAAA; ", width = "*", "GitHub")
    )
  , tags$tr(
      style = "align:top"
      , tags$td(style = "text-align:middle; background-color:#AAAAAA; "
                , map(.cran_libs[1:5], \(x) tags$li(x)) |> list() |> tags$ul())
      , tags$td(style = "text-align:middle; background-color:#AAAAAA; "
                , map(.cran_libs[6:length(.cran_libs)], \(x) tags$li(x)) |> list() |> tags$ul())
      , tags$td(style = "background-color:#EEEEEE; padding-right:20px; ", urls$git_libs)
      )
  ) |> tags$p()
```

## Data Summary <br><sup>Summarytools</sup> {.tabset .tabset-fade .tabset-pills}


```{r RETRIEVE_DATA, cache=TRUE,cache.lazy=TRUE, warning=FALSE, message=FALSE}

# MDRP database data
if (rlang::is_empty(dir(pattern = "mdrp_api_data[.][Rr]data$"))){
  if (!"api_data" %in% ls()){ 
    api_data <- urls$data |> 
      as.character() |> 
      stri_extract_all_regex("http.+csv", simplify = TRUE) |> 
      as.vector() %T>% 
      (\(x) tags$p(sprintf("Retrieve data from '%s'", x)) |> print())() |>
      GET() |>
      content() |>
      rawToChar() |>
      (\(x) read.csv(text = x, ))() |>
      as.data.table(na.rm = FALSE) |> 
      modify_at(c(4:7), as.character) %>% 
      modify_at(ls(., pattern = "Date"), lubridate::mdy) %>% 
      setnames(stri_replace_all_fixed(names(.) |> tolower(), ".", "_"))
  }
  
  
  if (!"api_dictionary" %in% ls()){ 
    api_dictionary <- invisible(urls$data |> 
      as.character() |> 
      stri_extract_all_regex("http.+pdf", simplify = TRUE) |> 
      as.vector() |>
      GET() |>
      content() |> 
      pdf_text())
    
    .summary_labels <- names(api_data) |> 
        rlang::set_names() |>
        map_chr(\(x) stri_replace_all_fixed(x, ".", " ", vectorize_all = FALSE)) |>
        imap_chr(\(x, y){ 
          api_dictionary |> 
            stri_extract_all_regex(
              sprintf(
                fmt = "(%s)[:]\n.+"
                , stri_replace_all_fixed(
                    x
                    , c("Pkg"
                        , "Intro"
                        , "COD Status"
                        , "FDA Application Number"
                        , "FDA Therapeutic Equivalence Code"
                        )
                    , c("Package"
                        , "Intro."
                        , "Covered Outpatient Drug [(]COD[)] Status"
                        , "FDA Application Number/OTC Monograph Number"
                        , "TEC"
                        )
                    , vectorize_all = FALSE
                    )
                )
              , simplify = TRUE
              ) |>
            stats::na.omit() |>
            as.vector() |>
            discard(\(x) x == "") %>%
            (\(i){ 
              .out <- ifelse(identical(character(), i), y, paste(i, collapse = "\n"))
              ifelse(stri_length(.out) > 50, paste0(.out, " ..."), .out)
            })()
        });
    
    iwalk(.summary_labels, \(x, y){ 
      .label = x; 
      api_data <<- modify_at(api_data, y, \(i){ attr(i, "label") <- .label; i }) 
    })
  }
  
} else { 
  dir(pattern = "mdrp_api_data") |> load() 
}

# openFDA supplementary data
if (rlang::is_empty(dir(pattern = "openFDA_ndc[.][Rr]data$"))){
  if (!"openFDA_ndc" %in% ls()){ openFDA_ndc <- (\(x){ 
      json.file <- "drug-ndc-0001-of-0001.json";
    
      if (!json.file %in% dir()){ 
        GET(urls$openFDA, write_disk(x, TRUE)) 
        unzip(zipfile = x)
      }
      
      read_json(path = json.file);
    })("ndc_json.zip") |> (\(x){
      data.table::rbindlist(modify_at(x$results, 1:length(x$results), as.data.table), fill = TRUE) |>
        setattr("metadata", x$meta)
    })()
  }
} else { 
  dir(pattern = "openFDA_ndc") |> load() 
}

```

```{r EXPLORATION, warning=FALSE, message=FALSE, fig.dim=c(600,960)}
api_data[runif(n= nrow(api_data)) <= 0.33, !"id"] |>
  dfSummary(labels.col = TRUE) |> 
  view(method = "render", report.title = "Medicaid Drug Rebate Program (MDRP) Data");

list(
  `OpenFDA: NDC Formats` = openFDA_ndc[, unique(product_ndc)] |> 
      sort() |> 
      stri_split_fixed("-") |> 
      map_chr(\(x) stri_length(x) |> paste(collapse = "-")) |> 
      table() |>
      ratio(type = "of.sum", decimals = 6)
  , `MDRP: NDC Formats` = api_data[, paste(stri_length(labeler_code), stri_length(product_code), sep = "-")] |>
      sort() |> 
      table() |>
      ratio(type = "of.sum", decimals = 6)
  ) |> 
  imap(\(x, y){ 
    tags$li(
      tags$b(y)
      , HTML(": <br>&nbsp;&nbsp; ")
      , sprintf("[%s]: %.2f %%", names(x), 100 * x) |> paste(collapse = "<br>&nbsp;&nbsp; ") |> HTML()
      )
    }) |>
  tags$ul() |>
  tags$p()
```

Create a function to check NDC formats and conform: 

```{r}
check_ndc_format <- \(lc, pc){ 
  pc <- modify_if(pc, \(i) stri_length(i) < 3, \(i) stri_pad_left(i, width = 3, pad = "0"))
  lc <- modify_if(
          lc
          , \(i) stri_length(i) < 4
          , \(i) stri_pad_left(i, width = ifelse(stri_length(pc) == 3, 5, 4), pad = "0")
          )
  
  paste(lc, pc, sep = "-")
}
```

This is done in order to join the data sets using a conformed NDC sequence.

```{r JOIN_DRUG_DATA, warning=FALSE, message=FALSE}
master_drug_data <- (\(x, i) x[
                  i, on = "alt_ndc==product_ndc"
                  , nomatch = 0
                  , allow.cartesian = TRUE
                  , `:=`(pharm_class = pharm_class, dea_schedule = dea_schedule, product_type = product_type, route = route, marketing_category = marketing_category)
                  , by = .EACHI
                  ])(api_data[, alt_ndc := map2_chr(labeler_code, product_code, check_ndc_format)]
                    , openFDA_ndc
                    )

master_drug_data %<>% modify_at("pharm_class", unlist)

```


## Temporal<br>Analysis 

```{r TIME_STUDY, warning=FALSE, message=FALSE}

if (rlang::is_empty(dir(pattern = "ndc_events[.][Rr]data$"))){ 
  if (!"ndc_events" %in% ls()){ 
    ndc_events <- define(
      api_data
      , ~ndc + fda_product_name + drug_category + fda_approval_date + 
          market_date + termination_date + reactivation_date
      , list(
          days_to_market = mean(market_date - fda_approval_date)
          , on_market_age = { 
              x <- ifelse(
                    is.na(reactivation_date)
                    , modify_if(termination_date, is.na, \(x) today()) - market_date 
                    , today() - reactivation_date
                    )
              mean(x)
            } 
          , days_market_absent = { 
              x <- ifelse(
                    is.na(termination_date)
                    , duration(0, "days")
                    , modify_if(reactivation_date, is.na, \(x) today()) - termination_date
                    )
              mean(x)
            } 
          ) ~ drug_category
      )
  }
  
} else {
  dir(pattern = "ndc_events[.][Rr]data$") |> load()
}

```

# Save Data

```{r SAVE_DATA}
  book.of.workflow::save_image("openFDA_ndc", file.name = "openFDA_ndc"
                               , use.prefix = FALSE, use.timestamp = FALSE, safe = FALSE)
  
  book.of.workflow::save_image("api_data", "api_dictionary", file.name = "mdrp_api_data"
                               , use.prefix = FALSE, use.timestamp = FALSE, safe = FALSE)
  
  book.of.workflow::save_image("ndc_events", file.name = "ndc_events"
                               , use.prefix = FALSE, use.timestamp = FALSE, safe = FALSE)

```