---
title: "Question I: Correlation"
output:
  html_notebook:
    css: markdown.css
    code_folding: hide
  html_document:
    df_print: paged
    css: markdown.css
    code_folding: hide
params:
  data_dir: data
  cran_libs: !r c('purrr', 'jsonlite', 'httr', 'summarytools', 'munsell', 'cachem', 'SmartEDA', 'htmltools', 'slider', 'stringi', 'magrittr', 'plotly', 'DT', 'data.table', 'pdftools', 'lubridate', 'future', 'furrr', 'future.callr')
  git_libs: !r paste0('book.of.', c('utilities', 'features', 'workflow')) |> c('architect', 'smart.data', 'event.vectors')
  refresh: !r FALSE
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  opts_chunk = list(cache=TRUE, cache.lazy=TRUE, warning=FALSE, message=FALSE)
  )

source("setup.R", local=TRUE)
```

```{=js}
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.7/MathJax.js?config=TeX-MML-AM_CHTML">
</script>
```

# {.tabset .tabset-fade .tabset-pills}

I decided to start asking some questions of the data given the metrics defined earlier.  I decided to look into the correlation between pairs of metrics relative to a third:

## Market Age 

"How does the correlation between *`days_to_market`* and *`on_market_age`* change by route of administration?":

```{r METRICS_CORRELATION_I} 
suppressWarnings(suppressMessages({
library(smart.data)

if (!"smrt_drugs" %in% .cache$keys()){
  smrt.drug_obs_data <- smart.data$
    new(x = .cache$get("drug_obs_data") |> print(), name = "drugs")$
    taxonomy.rule(
      term.map = if ("smrt_drug_taxonomy" %in% .cache$keys()){ .cache$get("smrt_drug_taxonomy") } else { data.table(term = "metrics", desc = "Metrics related to events and other descriptive statistics") }
      , update = !("smrt_drug_taxonomy" %in% .cache$keys())
      )$
    cache_mgr(action = upd) |>
    invisible();
  
  .cache$set("smrt_drugs", smrt.drug_obs_data);
  .cache$set("smrt_drug_taxonomy", smrt.drug_obs_data$smart.rules$for_usage);
} else { 
  invisible(smart.data$new(as.data.table(x = 1), "none"))
  (\(x){
    x$data <- .cache$get("drug_obs_data") |> unique()
    x$cache_mgr(action = upd)
    .cache$set("smrt_drugs", x);
  })(.cache$get("smrt_drugs"))
}

get.smart("drugs")$use(identifier, metrics, retain = c(route), subset = days_to_market >= 0) |> 
  # View() 
  setkey(route, alt_ndc, days_to_market, on_market_age) |>
  split_f(~route) |>
  map_dbl(\(x) x %$% cor(as.numeric(days_to_market), as.numeric(on_market_age))) |>
  modify_if(is.na, \(x) 0) |> 
  (\(x){ 
    x <- x[order(x)]
    nm <- names(x)
    z <- calc.zero_mean(x, as.zscore = TRUE, use.population = TRUE)
    y <- ratio(x + abs(min(x)), type="pareto", decimals = 6)
    
    .wh_scale <- 800 * c(1.2, .7)
    
    plot_ly(
      x = z
      , y = y
      , size = 5 * exp(x) + 10
      , width = .wh_scale[1]
      , height = .wh_scale[2]
      , hoverinfo = "text"
      , hovertext = sprintf(fmt ="<b>%s</b><br><b>Y:</b> %.2f%% of Total<br><b>Cor</b>(days_to_market, on_market_age): %.2f<br><b>Z-score</b>(X): %.2f", nm, y * 100, x, z)
      , color = x
      , stroke = I("black")
      , type = "scatter", mode = "markers"
      ) |>
      config(mathjax = "cdn", displayModeBar = FALSE) |>
      layout(
        xaxis = list(
            title = list(
              text = "Z-score<sub>X</sub>: X | Cor(m<sub>0</sub>, m<sub>1</sub>) ~ Route"
              , font = list(size = 16, family = "Georgia"))
            , gridcolor = "#FFFFFF"
            )
        , yaxis = list(
            title = list(
              text = "Cumulative Proportion (X)"
              , font = list(size = 16, family = "Georgia"))
            , gridcolor = "#FFFFFF"
            )
        , title = list(
            text = sprintf("Correlation Coefficient (<span style='text-emphasis-position:under; text-emphais: filled red double-circle; '>%s</span> vs. <span style='text-emphasis-position:under; text-emphais: filled red double-circle; '>%s</span>) by Route of Administration", "days_to_market", "on_market_age")
            , font = list(family = "Georgia"))
        , plot_bgcolor = rgb(.8,.8,.8)
        , margin = list(b = 30, t = 50)
        ) 
  })()
}))
```

## Days Absent from Market 

*"How does the correlation between `days_to_market` and `days_market_absent` change by route of administration?"*:

```{r METRICS_CORRELATION_II} 
get.smart("drugs")$use(identifier, metrics, retain = c(route), subset = days_to_market >= 0) |> 
  unique() |>
  setkey(route, alt_ndc, days_market_absent, on_market_age) |>
  split_f(~route) |>
  map_dbl(\(x) x %$% suppressWarnings(cor(as.numeric(days_market_absent), as.numeric(on_market_age)))) |>
  modify_if(is.na, \(x) 0) |> 
  (\(x){ 
    x <- x[order(x)]
    nm <- names(x)
    z <- calc.zero_mean(x, as.zscore = TRUE, use.population = TRUE)
    y <- ratio(x + abs(min(x)), type="pareto", decimals = 6)
    
    .wh_scale <- 800 * c(1.2, .7)
    
    plot_ly(
      x = z
      , y = y
      , size = 5 * exp(x) + 10
      , width = .wh_scale[1]
      , height = .wh_scale[2]
      , hoverinfo = "text"
      , hovertext = sprintf(fmt ="<b>%s</b><br><b>Y:</b> %.2f%% of Total<br><b>Cor</b>(days_to_market, days_market_absent): %.2f<br><b>Z-score</b>(X): %.2f", nm, y * 100, x, z)
      , color = x
      , stroke = I("black")
      , type = "scatter"
      , mode = "markers"
      ) |>
      config(mathjax = "cdn", displayModeBar = FALSE) |>
      layout(
        xaxis = list(
            title = list(
              text = "Z-score<sub>X</sub>: X | Cor(m<sub>0</sub>, m<sub>1</sub>) ~ Route"
              , font = list(size = 16, family = "Georgia"))
            , gridcolor = "#FFFFFF"
            )
        , yaxis = list(
            title = list(
              text = "Cumulative Proportion (X)"
              , font = list(size = 16, family = "Georgia"))
            , gridcolor = "#FFFFFF"
            )
        , title = list(
            text = sprintf("Correlation Coefficient (<span style='text-emphasis-position:under; text-emphais: filled red double-circle; '>%s</span> vs. <span style='text-emphasis-position:under; text-emphais: filled red double-circle; '>%s</span>) by Route of Administration", "days_to_market", "days_market_absent")
            , font = list(family = "Georgia"))
        , plot_bgcolor = rgb(.8,.8,.8)
        , margin = list(b = 30, t = 50)
        ) 
})()

```

</div>
Market age relative to days to market shows more variability in correlation distribution.  This is not to make an claim of statistically significant differentiation; however, it may be worth exploring whether or not there are clusters of administration routes based on event correlation.  A future update may address this, but this is as deep of exploration I'll go for now.
<br>
<br>
<div>

