---
title: "Question II: Event Network"
output:
  html_notebook:
    css: markdown.css
    code_folding: hide
  html_document:
    df_print: paged
    css: markdown.css
    code_folding: hide
params:
  data_dir: data
  cran_libs: !r c('purrr', 'jsonlite', 'httr', 'summarytools', 'munsell', 'cachem', 'SmartEDA', 'htmltools', 'slider', 'stringi', 'magrittr', 'plotly', 'DT', 'data.table', 'pdftools', 'lubridate', 'future', 'furrr', 'future.callr')
  git_libs: !r paste0('book.of.', c('utilities', 'features', 'workflow')) |> c('architect', 'smart.data', 'event.vectors')
  refresh: !r FALSE
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(opts_chunk = list(cache=TRUE, cache.lazy=TRUE, warning=FALSE, message=FALSE))
source("setup.R", local=TRUE)
```

```{=js}
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.7/MathJax.js?config=TeX-MML-AM_CHTML">
</script>
```

# {.tabset .tabset-fade .tabset-pills}

The next question I wanted to answer was if there were interesting event relations in the data. I decided to construct an event network and compare temporal proximity with key attributes of each event.  For the first pass of analysis, I stayed with the administration route (refered to as _route_ hereafter) as the grouping factor.  To create the temporal network, I used my custom package [`event.vectors`](https://github.com/delriaan/event.vectors) which leverages R6 classes.

## Preparation 

To prepare the input data, decisions about how to split the drug observation data into subgroups across which a common "join-key" is projected is required. 

> The use case for `event.vectors` is when you have more "join-key" values ($V$) expressed in orders of magnitude fewer event sources ($U: ||U|| \ll ||V||$)

Before investigation, I used a subset of the data conditioned on non-`NA` values for field `dea_schedule`. As a decision aid, I leveraged the [`dfSummary()`](mdd_summary.html) function from the [summarytools](https://www.rdocumentation.org/packages/summarytools/versions/1.0.1/topics/summarytools-package) package.

```{r PREP_SUMMARY, include=FALSE}
.subset_expr <- rlang::expr(
    (year(market_date) >= 2020) & 
    ((is.na(dea_schedule) & runif(length(dea_schedule)) < 0.5) | !is.na(dea_schedule))
    );
.summary_output <- summarytools::dfSummary(
    get.smart("drugs")$
      use(subset = !!.subset_expr) |> 
      unique() |>
      modify_at("dea_schedule", \(x) modify_if(x, is.na, \(x) "C0"))
    , display.labels = FALSE
    , graph.magnif = 0.5
    ) |> summarytools::view(method = "browser", file = "docs/mdd_summary.html")
```

After reviewing the [summary](docs/mdd_summary.html), I chose `dea_schedule` as the join and as the splitting factor used the first letter of values in `labeler_name` grouped into alphabetical sets of three or two.

```{r PREP_DATA}
# Split 'master_drug_data' and store in attached environment 'nb_env' ("notebook environment")
nb_env %$% rm(list = ls(pattern = "[A-Z]"));


letter_sets <- { slider::slide(
    LETTERS
    , .after = 2L
    , .step = 3
    , .f = \(x) sprintf("%s-%s", x[1], ifelse(is.na(x[3]), x[2], x[3]))
    ) |> 
    purrr::compact() |>
    unlist()
  }

define(
    get.smart("drugs")$use(subset=!!.subset_expr) |> unique()
    , modify_at(.SD, c("market_date", "termination_date"), \(x) as.Date(x, origin = origin)) |>
      modify_at("dea_schedule", \(x){ modify_if(x, is.na, \(i) "C0") })
    , .SD[!duplicated(alt_ndc)] ~ drug_category + dea_schedule
    , split_factor = pmap_chr({ 
          logic_map(
            fvec = stri_sub(labeler_name, length=1)
            , bvec = sprintf("[%s]", !!letter_sets)
            , regex = TRUE
            ) |> 
          as.data.table() %>% 
          setnames(stri_replace_all_regex(names(.), "\\[|(])", "", vectorise_all = FALSE))
        }, function(...) ...names()[c(...) == 1])
    , rel_key = 1
    , ~rel_key + split_factor + labeler_name + dea_schedule + drug_category + alt_ndc + fda_approval_date + market_date + termination_date + pharm_class + fda_product_name
    ) |> 
  unique() |>
  split(by = "split_factor") %T>% 
  # split(by = "dea_schedule") %T>% 
  (\(x) cat(sprintf("Created %s objects", length(x)), sep = "\n"))() |> 
  list2env(envir = nb_env)
```

## Sidebar {.tabset}

Next, I created the event network using `market_date` and `termination_date` to define event boundaries.  The significance of using this pair of temporal markers is that they are used to create a directed graph using _from_ &Rarr; _to_ sequencing.  

In addition, I defined a series of predicates to condition temporal cross-comparisons, ultimately resolved via Boolean *\&* having the global effect of reducing the number of edges across event networks:   

```{r EVS_PARAMS, echo=FALSE}
list(
  c("Predicate", "Impact")
  , c("mGap >= -?", "Only event pairs that overlap by no more than ? days are retained")
  , c("mGap <= ?", "Only event pairs that are separated by less than ? days are retained")
  , c("time.control = [0, ?]", "Only event pairs that span no more than ? days are retained<br>This is a global predicate that should be used with caution.")
  , c("unit = \"days\"", "Temporal fields will be forced to units of days during processing")
  ) |>
  imap(\(x, y){ 
    f <- if (y == 1){ tags$th } else { tags$td }
    tags$tr(
      f(class = "evs_params", x[1])
      , f(style="padding-left:10px;", HTML(x[2]))
      )
  }) |> 
  tags$table(id = "evs_params_table") 
```

To appreciate the need for such predication, it is worthwhile to take a moment to gain a sense for the computational space presented by this use case:

### Part I

Given event space $\mathbb{E}$ with $k$ independent event groupings: $$
\mathbb{E}:=\hat{e}_1 + \hat{e}_2+  \hat{e}_3 + \cdots + \hat{e}_k
$$
and corresponding heterogeneous data tables ($\mathbb{D} :=D_1,D_2,D_3,\cdots,D_k$) each having a "join-key" ($J$, a column or column tuple shared across $\mathbb{D}$), an event vector ($\overrightarrow{V_q}$) is formed having values indicating the existence or absence of selection over $D_k$ predicated on $q$ unique values $\forall J$: 

$$
\overrightarrow{V_q}:= 
\exists \sigma_{j_q \in J}(D_k) \Rightarrow
\begin{pmatrix}0\\1\end{pmatrix}_{q}\hat{e}_1 + 
\begin{pmatrix}0\\1\end{pmatrix}_{q}\hat{e}_2 + 
\begin{pmatrix}0\\1\end{pmatrix}_{q}\hat{e}_3 + 
\cdots + 
\begin{pmatrix}0\\1\end{pmatrix}_{q}\hat{e}_k \\
q :=\text{ the q}^{th}\text{ distinct value in }J\enspace \forall \mathbb{D} \\
\sigma_{j_q \in J}(D_k) \equiv \dot{D_{qk}}:=\text{ predicated selection operation in a relational algebra context}
$$


From each $\overrightarrow{V_q}$, $q$ subsets over $\mathbb{D}$ are derived as $\Gamma_q:=D_{k_q}|\overrightarrow{V_q}\ne\emptyset$. The number of records ($n$) resulting from selection across the data sets can be represented as an array of row vectors: 

$$
\Gamma_q\longrightarrow \text{Distinct row counts over}\dot{D_{k_q}}\Rightarrow\begin{array}
((n_{11}&n_{12}&n_{13}&\cdots&n_{1k})\\
(n_{21}&n_{22}&n_{23}&\cdots&n_{2k})\\
(n_{31}&n_{32}&n_{33}&\cdots&n_{3k})\\
\vdots&\vdots&\vdots&\vdots&\vdots\\
(n_{q1}&n_{q2}&n_{q3}&\cdots&n_{qk})
\end{array}
$$

### Part II

Next, considering a [fully-connected](https://en.wikipedia.org/wiki/Complete_graph) graph having $\beta$ nodes and number of edges given as follows: $$
\begin{align}
K\rightarrow f(\beta) := {\beta\choose2} \equiv \frac{\beta(\beta-1)}{2}
\end{align}
$$
 and taking into account that `event.vectors` creates separate graphs for each $q$, the number of edges per fully-connected graph $G_q$ is given as follows: 

$$
\beta_q := \sum_{i=1}^q\Gamma_i \Rightarrow (\beta_1,\beta_2,\beta_3,\cdots,\beta_q)\\
K_q \rightarrow f(\beta_q) := {\beta_q\choose2}
$$
, which is also the number of comparisons to be carried out computationally.

$K_q$ can be reduced by projecting a set of logically-reduced predicates (as mentioned earlier) to interim results before creating each of the $q$ graphs:

$$
P_m:m\text{ predicates}\Rightarrow
\begin{pmatrix}0\\1\end{pmatrix}_m\\
\dot\beta_q:=\forall i \in q\rightarrow\Gamma_i*\Pi (P_m|\dot D_{k_i})\\ 0 \le \dot\beta_q\le\beta_q
$$ 
This can quickly result in not only edge reduction but graph elimination for each $q$ ($\Gamma_q$ can easily become sparse).  Trial and error may needed to fine-tune how graphs are generated as well as improve execution performance which is why the `graph.control` argument is provided for `event.vectors::make.evs_universe()` to allow for operating on graphs after the fact.

## Execution and Wrap Up 

The following parameters were used to generate event graphs:

- `mGap >= -60`
- `mGap <= 90`
- `time.control = list(0, 181)`
- `unit = "days"`
 
and packages `future` and `furrr` were invoked to enable parallelism (externally and internally, respectively).

    
```{r EVENT_NETWORK}
load_unloaded(event.vectors, furrr, tictoc, igraph);

mdrp_events <- event.vectors$new();

mdrp_events$configure(
  src.defs = sprintf("nb_env[[\"%s\"]]", ls("nb_env", pattern = "[A-Z]")) |> rlang::parse_exprs()
  , contexts = ls("nb_env", pattern = "[A-Z]")
  , map.fields = {
      replicate(
        n = ls("nb_env", pattern = "[A-Z]") |> uniqueN()
        , expr = { c("dea_schedule", "market_date", "termination_date")}
        , simplify = FALSE
        )
      }
  , chatty = !TRUE
  );

future_cluster <- makeClusterPSOCK(5);
plan(tweak(cluster, workers = future_cluster));

tic(glue::glue("{Sys.time()} Creating the Event Vectors"));
mdrp_events$
  make.evs_universe(
    mGap >= -60
    , mGap <= 90
    , time.control = list(0, 181)
    , unit = "days"
    , furrr_opts = furrr::furrr_options(scheduling = Inf, seed = TRUE)
    , graph.only = FALSE
    , chatty = FALSE
    );
toc(log = TRUE);

suppressMessages({
  plan(sequential)
  if ("future_cluster" %in% ls()){ 
    parallel::stopCluster(future_cluster)
    rm(future_cluster)
  }
  invisible(gc())
});
```

After generating the universe of events I explored the graphs using [`visNetwork`](https://visjs.github.io/vis-network/docs/network/). Only one graph was interesting enough to retain which can be viewed [here](./viznetwork_C0.html).

```{r EVENT_NETWORK_VIZ, cache=FALSE}
f2ab <- list(theta = 0.1, gravitationalConstant = -500, centralGravity = 0.0,  avoidOverlap = 1, damping = 0.7);
viz_file <- "./viznetwork_%s.html"

#
ev_plot <- \(g){
  .css <- "background-color:#444444; color:#DDDDDD; border-radius:10; font-weight:bold; "
  
  # Vertices (https://rdrr.io/cran/visNetwork/man/visNodes.html) ----
  V(g)$trace <- retrace.evs(event_graph = g, evs = mdrp_events);
  
  g0 <- V(g)$name |> map(\(x){ 
      stri_replace_first_regex(x, pattern = "[:][0-9]+", replacement = "") |> 
        # stri_split_fixed(" | ", simplify = TRUE) |> 
        unlist()
        paste(collapse = ",")
    })
  
  g1 <- V(g)$trace |> map(\(x) x$labeler_name |> unique() |> paste(collapse = ","))
  
  g2 <- sprintf("In-Degree Quantile Band: %s", degree(graph = g, mode = "in", loops = FALSE) |> make.quantiles(as.factor = TRUE, 0:10/10) |> as.character())
  
  V(g)$group <- paste(g0, g1, g2, sep=",")
  
  V(g)$title <- map(V(g)$trace, \(x){
    sprintf(
      fmt = "<b>%s</b><br><ul>%s</ul>"
      , x$labeler_name
      , sprintf(
          "<li>%s</li>"
          , x[, sprintf(
                  "%s days on market (%s)"
                  , x$on_market_age
                  , stri_extract_first_regex(x$on_market_age |> as.numeric(unit = "days") |> ddays(), "[~].+s")
                  ) |> 
                c(alt_ndc, fda_product_name)
              ]
          ) |> paste(collapse = "")
      )
    })
    
  
	V(g)$color <- degree(graph = g, mode = "total", loops = FALSE) |>
          		  ratio(type = "of.max", decimals = 6) |> 
          		  rgb(red = 0.500000, blue = 0.500000)
  
  # Edges (https://rdrr.io/cran/visNetwork/man/visEdges.html) ----
  E(g)$title <- edge.attributes(g) %$% {
    src.pair |> 
      stri_split_fixed(" -> ", simplify = TRUE) |> 
      (\(x) list(src_from = x[, 1], src_to = x[, 2]))() |> 
      list2env(envir = environment())
    
    .css <- "background-color:#DDDDDDEE; border:solid 1px #CCCCCCEE; border-radius:10; font-weight:bold; width:120%; padding-top:10px; padding-bottom:10px; "
    sprintf(
      paste0(
        "<p style=\"%s\"><b style=\"font-size:larger; \">&nbsp;&nbsp;%s <span style=\"font-family:Georgia\">(&epsilon; = %s)</span></b>"
        , "<br><br>&nbsp;<u>From</u> {%s} <sup><%s></sup>"
        , "<br><br>&nbsp;<u>To</u> {%s} <sup><%s></sup></p>"
        , "</p>"
        )
      , .css, epsilon.desc, round(epsilon, 4) |> as.character()
      , src_from, from.coord
      , src_to, to.coord
      )
  }
	
	E(g)$smooth <- TRUE
	
	E(g)$length <- map(E(g)$mGap, \(x) 50 * ratio((x - abs(min(x))), type = "of.max", decimals = 4))
	
	E(g)$dashes <- E(g)$mGap |> unlist() > 0
	
	E(g)$color <- map(E(g), \(x){ 
	    v_ends.color <- ends(graph = g, es = x) |> map(\(i) V(g)[[i]]$color |> colorspace::hex2RGB()) 
	    list(color = colorspace::mixcolor(
	              alpha = 0.5
	              , color1 = v_ends.color[[1]]
	              , color2 = v_ends.color[[2]]
	              ) |> colorspace::hex()
	         , opacity = ifelse(x$mGap <= 0, 1, 0.5)
	         )
	  })
	
	# Update & Output ----
	.vizWidth = 1400
	gph <<- g
	visIgraph(igraph = g, physics = TRUE, type = "full") |>
  	visPhysics(solver = "forceAtlas2Based", timestep = 0.05) |>
  	visOptions(width = .vizWidth, height = (12/16) * .vizWidth, selectedBy = list(variable = "group", multiple = TRUE)) |> 
	  visEdges(smooth = FALSE) |>
	  visNodes(font = list(face = "Georgia", size="40", color = "black", background = "#DEE", bold=TRUE))
}

#
mdrp_events$evt_graphs |> 
  iwalk(\(x, y){ 
    ev_plot(x) |> 
      htmlwidgets::prependContent(
        tags$table(
          tags$tr(
              tags$td(style='font-weight:bold; align:right; padding-right:10px', "Edges:")
              , tags$td("Dashes indicate non-overlapping event durations")
              )
          , tags$tr(
              tags$td(style='font-weight:bold; align:right; padding-right:10px', "Colors: ")
              , tags$td(HTML("Based on relative proportionality of <code>degree</code> vertex metrics in the graph"))
              )
          , tags$tr(
              tags$td(style='font-weight:bold; align:right; padding-right:10px', "Drop-down menu:")
              , tags$td("Use to select predefined vertex groups")
              )
          )
        ) |>
      saveWidget(file = sprintf(viz_file, y)) 
});

```

From here, one could conduct some graph analysis or graph-based predictive modeling, but for now, I will close this exploratory exercise.  I hope you have found something interesting about the methods employed, packages used, and results.  If you would like to follow up with me, I can be reached via [LinkedIn](https://www.linkedin.com/in/chionesu-george-1067b122/).

&mdash; Thank you!

<br>
<br>