---
title: "Question II: Event Clusters"
output:
  html_notebook:
    css: markdown.css
    code_folding: hide
  html_document:
    df_print: paged
    css: markdown.css
    code_folding: hide
params:
  data_dir: data
  cran_libs: !r c('purrr', 'jsonlite', 'httr', 'summarytools', 'munsell', 'cachem', 'SmartEDA', 'htmltools', 'slider', 'stringi', 'magrittr', 'plotly', 'DT', 'data.table', 'pdftools', 'lubridate', 'future', 'furrr', 'future.callr')
  git_libs: !r paste0('book.of.', c('utilities', 'features', 'workflow')) |> c('architect', 'smart.data', 'event.vectors')
  refresh: !r FALSE
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(opts_chunk = list(cache=TRUE, cache.lazy=TRUE, warning=FALSE, message=FALSE))
source("setup.R", local=TRUE)
```

```{=js}
<script type="text/javascript" async
  src="https://cdnjs.cloudflare.com/ajax/libs/mathjax/2.7.7/MathJax.js?config=TeX-MML-AM_CHTML">
</script>
```

# {.tabset .tabset-fade .tabset-pills}

The next question I wanted to answer was if there were meaningful event clusters in the data. I decided to construct an event network and compare temporal proximity with key attributes of each event.  For the first pass of analysis, I stayed with the administration route (refered to as _route_ hereafter) as the grouping factor.  To create the temporal network, I used my custom package [`event.vectors`](https://github.com/delriaan/event.vectors) which leverages R6 classes.

## Preparation 

To prepare the input data, I had to decide how to split `master_drug_data` into smaller data tables based on one or more shared features (`event.vectors` requires at least two classes of data for cross-comparison).  As a decision aid, I leveraged the [`dfSummary()`](mdd_summary.html) function from the [summarytools](https://www.rdocumentation.org/packages/summarytools/versions/1.0.1/topics/summarytools-package) package.

```{r PREP_SUMMARY}
summarytools::dfSummary(
  master_drug_data
  , display.labels = FALSE
  , graph.magnif = 0.5
  ) |> 
  summarytools::view(
    method = "browser"
    , file = "docs/mdd_summary.html"
    )
```


After reviewing the summary, choosing one of `cod_status`, `drug_category` or `dea_schedule` made the most sense as the number of levels were not too many given the scale of comparisons considered (${n \choose k} + {n\choose k}^* \equiv 2{n\choose k}$, where ${n\choose k}^*$ is the mirror of ${n\choose k}$).

In preparing to use the `event.vectors` package, decisions about how to split the drug observation data into subgroups across which a common "join-key" is projected is required. As a heuristic in checking for sufficient interaction counts, called $\text{replications(formula, ...)}_i$: $i$ calls to to `replications()` using as the formula argument $$\text{formula}:=\enspace \sim \text{cod_status}*\text{drug_category} + \text{cod_status}*\text{dea_schedule} + \text{drug_category}:\text{dea_schedule}$$

```{r PREP_FREQ_TBL}
mdrp_crosstab <- replicate(n = 30, { 
  replications(
    formula = ~ cod_status*drug_category + cod_status*dea_schedule + drug_category:dea_schedule
    , data = get.smart("drugs")$
              use(identifier
                  , retain = c(dea, drug_cat, cod_stat)
                  , subset = runif(n = length(alt_ndc)) <= 0.25
                  ) |> modify_at("cod_status", as.factor)
    )
  })

setattr(
  mdrp_crosstab
  , "interactions"
  , apply(mdrp_crosstab
      , MARGIN = 1
      , FUN = \(k) reduce(k, rbind) |> apply(2, \(x) Rmisc::CI(x))
      ) |> print()
  )

```

The data will need to be filtered for those levels of `cod_status` returning *0*. I decided to use `drug_category` and `dea_schedule` to split drug observation data with `cod_status` as the "join-key" (see `?``event.vectors-package`` `).  

Also, I limited the data to market dates starting in 2019 and forced all values in `termination_date` more than six months in the future (relative to "today") to six months in the future: this was done in order to preserve the sense of "the future" without unnecessarily allowing forlarge temporal differences in light of the way `event.vectors` works.

```{r PREP_DATA}
# Split 'master_drug_data' and store in attached environment 'nb_env' ("notebook environment")
nb_env %$% rm(list = ls(pattern = "^C"))

get.smart("drugs")$
  use(subset = (year(market_date) >= 2019) & !(cod_status %in% c(6:9, 11))) |>
  define(
    rel_key = paste0("mkt_dt_qtr_", quarter(market_date))
    , modify_at(.SD, c("market_date", "termination_date"), \(x) as.Date(x, origin = origin)) 
    , .SD[!duplicated(alt_ndc)] ~ drug_category + dea_schedule
    ) |> 
  split(by = c("drug_category", "dea_schedule")) %T>% 
  (\(x) cat(sprintf("Created %s objects", length(x)), sep = "\n"))() %>% 
  rlang::set_names(names(.) |> stri_replace_first_fixed("NA", "C0")) |>
  list2env(envir = nb_env)
```

## Execution

Next, I created the event network, using `market_date` and `termination_date` to define and event duration.  The significance of using this pair of temporal markers is that they are used to create a directed graph using _from_ &Rarr; _to_ sequencing:

```{r EVENT_NETWORK}
load_unloaded(event.vectors, furrr, tictoc, igraph)

mdrp_events <- event.vectors$new()

mdrp_events$configure(
  src.defs = paste("nb_env", ls("nb_env", pattern = "^[A-Z]"), sep = " %$% ") |> rlang::parse_exprs()
  , contexts = paste0("DrugCat: ", ls("nb_env", pattern = "^[A-Z]") |> stri_replace_first_fixed(".", " | Sched: "))
  , map.fields = {
      replicate(
        n = ls("nb_env", pattern = "^[A-Z]") |> uniqueN()
        , expr = { c("rel_key", "market_date", "termination_date")}
        , simplify = FALSE
        )
      }
  , chatty = !TRUE
  )

tic(glue::glue("{Sys.time()} Creating the Event Vectors"))
future_cluster <- makeClusterPSOCK(5)
plan(tweak(cluster, workers = future_cluster))

mdrp_events$
  make.evs_universe(
    mGap >= -30
    , mGap <= 30
    , mSt >= 90
    , time.control = list(0, 181)
    , unit = "days"
    , furrr_opts = furrr::furrr_options(scheduling = Inf, seed = TRUE)
    , graph.only = FALSE
    , chatty = FALSE
    )
toc(log = TRUE)

suppressMessages({
  plan(sequential)
  if ("future_cluster" %in% ls()){ 
    parallel::stopCluster(future_cluster)
    rm(future_cluster)
  }
  invisible(gc())
})
```

I then randomly selected an event network to visualize using [`visNetwork`](https://visjs.github.io/vis-network/docs/network/).  The output is located [here](viznetwork.html)

```{r EVENT_NETWORK_VIZ, cache=FALSE}
f2ab <- list(theta = 0.1, gravitationalConstant = -500, centralGravity = 0.0,  avoidOverlap = 1, damping = 0.7);
gph <- mdrp_events$evt_graphs[sample(names(mdrp_events$evt_graphs), 1)]
viz_file <- "./docs/viznetwork.html"

gph_trace <- retrace.evs(event_graph = gph[[1]], evs = mdrp_events)

ev_plot <- gph[[1]] |> (\(g){
    .css <- "background-color:#444444; color:#DDDDDD; border-radius:10; font-weight:bold; "
    
    .table_plotly <- plotly::plot_ly()
    
    igraph::V(g)$trace <- gph_trace;
    igraph::V(g)$title <- map(gph_trace, \(x){ 
      sprintf("<b>Market Age</b>: %s days (%s)<h5>%s</h5>"
              , x$on_market_age
              , stri_extract_first_regex(x$on_market_age |> as.numeric(unit = "days") |> ddays(), "[~].+s")
              , x$labeler_name
              )
    })
    
    igraph::E(g)$title <- igraph::edge.attributes(g) %$% {
      src.pair |> 
        stri_replace_all_regex("[:][0-9]+", "", vectorize_all = FALSE) |>
        as.character() |>
        stri_split_fixed(" -> ", simplify = TRUE) |> 
        (\(x) list(src_from = x[, 1], src_to = x[, 2]))() |> 
        list2env(envir = environment()) 
      
      .css <- "background-color:#444444; color:#DDDDDD; border-radius:10; font-weight:bold; width:20%; "
      
      sprintf(
        paste(
          "<p style=\"%s\"><b style=\"font-size:larger; \">%s (%s) </b>"
          , "<u>From</u> {%s} <sup><%s></sup>"
          , "<u>To</u> {%s} <sup><%s></sup></p>"
          , sep = "<br><br>"
          )
        , .css, epsilon.desc, mGap, src_from, from.coord, src_to, to.coord
        ) |>
      # stri_replace_all_fixed("@", ", ", vectorize_all = FALSE) |>
      map(\(x) HTML(x))
    }
      
		igraph::V(g)$color <- igraph::V(g)$trace |> 
		    map_dbl(\(x) x$market_date) |> 
		    (\(x) x - min(x))() |>
  		  ratio(type = "pareto", decimals = 6) |> 
  		  rgb(red = 0.500000, blue = 0.500000)
		
		igraph::E(g)$smooth <- TRUE
		
		g
	})() |>
	visNetwork::visIgraph(physics = TRUE, type = "full") |>
	visNetwork::visPhysics(solver = "forceAtlas2Based", timestep = 0.05) |>
	visNetwork::visOptions(width = "1600", height = "1024") |>
  visNetwork::visEdges(shadow = TRUE) %T>%
  saveWidget(file = viz_file)

# browseURL("docs/viznetwork.html")
```
